---
title: "Preliminary Analyses of DEAP"
output: html_document
date: "2024-04-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages
The main two packages used for this are the lme4 package and the flexplot 
package. lme4 is well known for it's use in linear modeling. Flexplot is a less
well known package created by Dr. Dustin Fife, Professor of Statistics at Rowan
University in New Jersey. He created the packages as a tool for visualizing and
comparing the fit of different models. The other two packages are not specific
to this analysis.

## Analysis
Import Packages
```{r Import packages, eval=FALSE}
library(tidyverse)
library(party)
library(lme4)
```


```{r}
  library(tidyverse)
library(flexplot)
library(party)
library(lme4)

# Normalize function
normalize <- function(input) {
  avg = mean(input)
  std = sd(input)
  return((input - avg) / std)
}

# Set wd to active script location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# Read in  CSV files (Change these paths to whatever they need to be)
physio_formatted <- read_csv("../data/deap/features/physio-formatted.csv")
mir_formatted <- read_csv("../data/deap/features/mir-formatted.csv")

# Select physio column of interest (Look at the physio_formatted data.frame for
# Available column names)
outcome_variable <- "gsr_mean"

# normalize gsr_mean (only necessary if outcome is 'gsr_mean', but doesn't hurt
# regardless)
physio_formatted$gsr_mean = log(physio_formatted$gsr_mean)

# Create 3 column data frame with column of interest
reduced_physio <- physio_formatted |> 
  select(c("participant", "stim_id", outcome_variable)) |> 
  rename(outcome = outcome_variable) |> 
  filter(participant != "18")

flexplot(outcome ~ 1, data=reduced_physio)

# Create Model data
model_data <- reduced_physio |> 
  left_join(mir_formatted, by = "stim_id")

# Model 1: Std dev of Zero Crossing rate as a fixed effect vs data mean
full <- lmer(outcome ~ zero_cross_std + (1|participant), data = model_data)
reduced <- lmer(outcome ~ 1 + (1|participant), data = model_data)
compare.fits(outcome ~ zero_cross_std, data=model_data, full, reduced)
model.comparison(full, reduced)
visualize(full)

# Model 2: Std dev of Zero Crossing rate as fixed and random effect vs only as fixed effect (ie fixed slopes vs variable slopes)
full <- lmer(outcome ~ zero_cross_std + (zero_cross_std|participant), data = model_data)
reduced <- lmer(outcome ~ zero_cross_std + (1|participant), data = model_data)
compare.fits(outcome ~ zero_cross_std, data=model_data, full, reduced)
model.comparison(full, reduced)
visualize(full)

# Model 3: Std Dev of MFCC 6 as a fixed effect vs data mean
full <- lmer(outcome ~ mfcc6_std + (1|participant), data = model_data)
reduced <- lmer(outcome ~ 1 + (1|participant), data = model_data)
compare.fits(outcome ~ mfcc6_std, data=model_data, full, reduced)
model.comparison(full, reduced)
visualize(full)

# Model 4: Std Dev of MFCC 6 as a fixed effect and random effect vs only as fixed effect (ie fixed slopes vs variable slopes)
full <- lmer(outcome ~ mfcc6_std + (mfcc6_std|participant), data = model_data)
reduced <- lmer(outcome ~ mfcc6_std + (1|participant), data = model_data)
compare.fits(outcome ~ mfcc6_std, data=model_data, full, reduced)
model.comparison(full, reduced)
visualize(full)

# For GSR RMS
full <- lmer(outcome ~ zero_cross_std + mfcc6_std + (zero_cross_std|participant), data = model_data)
reduced <- lmer(outcome ~ zero_cross_std + (zero_cross_std|participant), data=model_data)
compare.fits(outcome ~ zero_cross_std, data=model_data, full, reduced)
model.comparison(full, reduced)
visualize(full, formula = outcome ~ mfcc6_std | zero_cross_std)
```

